---
title: "TICS subscore and Langa-Weir Cognition Exploration"
author: "Ruby Hickman"
date: "2/13/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7)
```

```{r, include = FALSE}
library(ggplot2)
library(epiDisplay)
library(dplyr)
library(corrplot)

load("/Users/rubyhickman/Documents/01_michigan/work/cognition_mental_work/danny work/df_sub_04.rda")
```

Exploring distributions of the "langa" variable, which is actually just the raw TICS subscore (0-27) which
goes into calculating the langa-weir cognition categories.  
```{r}
#tics subscore of entire sample 
ggplot(df.sub) + 
  aes(x = df.sub$langa) +
  labs(x = "TICS subscore (0-27), unstratified (n=2774)") +
  geom_histogram(color = "#CFBAE1", fill = "#D6CCE0", binwidth = 1)

#stratified by physical work demands
ggplot(df.sub) + 
  aes(x = df.sub$langa, fill = df.sub$physical.cut) +
  labs(x = "TICS subscore (0-27), stratified by low vs. high physical demands", fill = "Mean physical \n demand score") +
  geom_histogram(color = "#C59FC9", alpha = 0.7, binwidth = 1, position = 'identity')
t.test(df.sub$langa ~ df.sub$physical.cut) #significant difference 

#mental work demands 
df.sub$md_cut <- cut(df.sub$mental.demand.average, quantile(df.sub$mental.demand.average, c(0, 0.5, 1.0)), include.lowest = T)
ggplot(df.sub) + 
  aes(x = df.sub$langa, fill = df.sub$md_cut) +
  labs(x = "TICS subscore (0-27), stratified by low vs. high mental demands", fill = "Mean mental \n demand score") +
  geom_histogram(color = "#C59FC9", alpha = 0.7, binwidth = 1, position = 'identity')
t.test(df.sub$langa ~ df.sub$md_cut) #significant difference 

#volunteering/helping friends 
ggplot(df.sub) + 
  aes(x = df.sub$langa, fill = factor(df.sub$volunt.help.bin)) +
  labs(x = "TICS subscore (0-27), stratified by volunteering/helping friends", fill = "Volunteering or helping \n friends (1 = yes?) \n ≥50hr/yr?") +
  geom_histogram(color = "#C59FC9", alpha = 0.7, binwidth = 1, position = 'identity')
t.test(df.sub$langa ~ df.sub$volunt.help.bin) #significant difference 

#practice effect 
ggplot(df.sub) + 
  aes(x = df.sub$langa, fill = as.factor(df.sub$prac.effect)) +
  labs(x = "TICS subscore (0-27), stratified by practice effect", fill = "Practice effect \n 1 = Yes") +
  geom_histogram(color = "#C59FC9", alpha = 0.7, binwidth = 1, position = 'identity')
t.test(langa ~ prac.effect, data = df.sub) #not a significant difference in means
```
  
To analyze whether langa-weir is an acceptable option, also categorized the TICS subscores into (actual) langa weir scores to analyze by groups.  
```{r}
#create categorical langa-weir category 
df.sub <- mutate(df.sub, langa_weir = case_when(langa < 7 ~ "dementia", 
                                      langa >= 7 & langa < 12 ~ "CIND", 
                                      langa <=27 ~ "normal")) 
#now dichotomous normal vs. impaired cognition 
df.sub <- mutate(df.sub, langa_weir_dic = case_when(langa < 12 ~ "impaired", 
                                      langa <=27 ~ "normal")) 
df.sub$langa_weir <- factor(df.sub$langa_weir, levels = c("dementia", "CIND", "normal"))

#all 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic) +
  labs(x = "Langa-Weir Categorizations") +
  geom_bar(fill = "#ACACDE", col = "#6E6E8E") 

#by physical work demands 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = df.sub$physical.cut) +
  labs(x = "Langa-Weir categorizations by physical work demands", fill = "Mean physical \n demand score") +
  geom_bar(col = "#C59FC9", position = "dodge")
chisq.test(df.sub$langa_weir, df.sub$physical.cut)
  #proportions of those in each outcome who had the exposure 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = df.sub$physical.cut) +
  labs(x = "Proportions of high physical demand status by cognitive status", fill = "Mean physical \n demand score") +
  geom_bar(col = "#C59FC9", position = "fill")
chisq.test(df.sub$langa_weir_dic, df.sub$physical.cut)

#by mental work demands 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = df.sub$md_cut) +
  labs(x = "Langa-Weir categorizations by mental work demands", fill = "Mean mental \n demand score") +
  geom_bar(col = "#C59FC9", position = "dodge")
chisq.test(df.sub$langa_weir, df.sub$md_cut)
  #proportions of those in each outcome who had the exposure 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = df.sub$md_cut) +
  labs(x = "Proportions of high mental demand status by cognitive status", fill = "Mean mental \n demand score") +
  geom_bar(col = "#C59FC9", position = "fill")
chisq.test(df.sub$langa_weir_dic, df.sub$md_cut)

#by volunteering/helping friends: total counts 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = factor(df.sub$volunt.help.bin)) +
  labs(x = "Langa-Weir categorizations by volunteering/helping friends", fill = "Volunteering or helping \n friends (1 = yes?) \n ≥50hr/yr?") +
  geom_bar(col = "#C59FC9", position = "dodge")
chisq.test(df.sub$langa_weir, df.sub$volunt.help.bin)
  #proportions of those in each outcome who had the exposure 
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = factor(df.sub$volunt.help.bin)) +
  labs(x = "Proportions of volunteering status by cognitive status", fill = "Volunteering or helping \n friends (1 = yes?) \n ≥50hr/yr?") +
  geom_bar(col = "#C59FC9", position = "fill")
chisq.test(df.sub$langa_weir_dic, df.sub$volunt.help.bin)

#by practice effect  
ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = as.factor(df.sub$prac.effect)) +
  labs(x = "Langa-Weir categorizations stratified by practice effect", fill = "Practice effect \n 1: Yes") +
  geom_bar(col = "#C59FC9", position = "dodge")
chisq.test(df.sub$langa_weir, df.sub$prac.effect)

ggplot(data = df.sub) + 
  aes(x = langa_weir_dic, fill = as.factor(df.sub$prac.effect)) +
  labs(x = "Proportions of practice effect status by cognitive status", fill = "Practice effect \n 1: Yes") +
  geom_bar(col = "#C59FC9", position = "dodge")
chisq.test(df.sub$langa_weir_dic, df.sub$prac.effect) #not a significant p-value 
```
  
Relationship between exposure variables -- are they related in the ways we would expect?  
```{r}
#Mental work demands and physical work demands 
cor(df.sub$avg.physical.job, df.sub$mental.demand.average)

ggplot(df.sub) + aes(x = df.sub$avg.physical.job, y = df.sub$mental.demand.average) + 
  geom_point(shape = 1, col = "#8367C7") + 
  geom_smooth(col = "#053225", method = lm) + 
  labs(x = "Mean Physical Work Demands", y = "Mean Mental Work Demands")

#Mental work demands and volunteering/helping friends 
ggplot(df.sub, aes(factor(volunt.help.bin), y = mental.demand.average)) + 
  geom_boxplot(fill = c("#E34A6F", "#BB7E8C")) + 
  labs(x = "Volunteering or helping friends (1 = yes?) ≥50hr/yr", y = "Mean Mental Work Demands")
t.test(df.sub$mental.demand.average ~ df.sub$volunt.help.bin) #significant 

#Mean physical work demands and volunteering/helping friends 
ggplot(df.sub, aes(factor(volunt.help.bin), y = avg.physical.job)) + 
  geom_boxplot(fill = c("#E34A6F", "#BB7E8C")) + 
  labs(x = "Volunteering or helping friends (1 = yes?) ≥50hr/yr", y = "Mean Physical Work Demands")
t.test(df.sub$avg.physical.job ~ df.sub$volunt.help.bin) #not significant 
```

